\begin{acknowledgements}
  % General intro sentence
  % The completion of this thesis would not have been possible without all support 
  % Nicole

  This thesis was not completed in a vacuum, and as such there are many people I wish to thank for the help and support they provided throughout my candidature.

  First and foremost, I want to thank my supervisor, Nicole Bell, for all the wisdom, guidance, and patience you have provided me over the last five years. Studying under your supervision has been a wonderful experience, and I cannot thank you enough for giving me this opportunity.

  % Collaborators
  I was fortunate enough to collaborate with Sandra Robles and Giorgio Busoni on many of the projects I worked on during my PhD. Thank you both for all that you have taught me over these years. In particular, I want to thank Sandra for always pushing me to do the best work possible. Your mentorship was truly appreciated.

  % Panel
  To the remainng members of my advisory panel, Matthew Dolan and Andrew Melatos, thank you for the guidance and encouragement you have provided over the years.

  % Office mates
  Returning to the office after nearly two years of COVID lockdowns solidified how important working alongside fellow students is to the PhD experience. In particular, I would like to thank Fred Hiskens, Maaz Hayat, Ayo Ore, Alex Ritter, Alexei Sopov, and Josh Wood for the countless hours of conversation we shared, both physics and non-physics related. You all made working in that office a truley wonderful experience.
 
  % Friends
  I was lucky enough to meet and become good friends with many people in and around the physics department throughout my long tenure at this university. In particular, I would like to thank Marcel Hohmann, Mike Mews, Oliver Anagnostou, Daniel Marcantonio, Lachlan Milligan, and Neuton Li. Your friendship over these long years has been truly appreciated.

  % Parents and family
  Finally, I am incredibly thankful to my family for all the support they have provided me while completing this degree. Thank you for the love and encouragement you have provided over the years. 

  % Acknowledgement of country
  \vfill
  \noindent\rule{\textwidth}{0.5pt}
  \textit{I would like to acknowledge the traditional owners and custodians of the lands on which this work was completed, the Wurundjeri people of the Kulin nation, and pay my respects to Indigenous Elders past and present. Sovereignty has never been ceded. It is, was, and always will be, Aboriginal land.}
\end{acknowledgements}
