\begin{preface}


This thesis is comprised of seven chapters, with chapters~\ref{chapter:capture_intro} -~\ref{chapter:thermalisation} based on several of the papers produced during the completion of this degree, namely Refs.~\cite{Bell:2020jou_sep_ImprovedTreatmentDark, Bell:2020lmm_mar_ImprovedTreatmentDark, Bell:2021fye_oct_Improvedtreatmentdark, Anzuini:2021lnv_nov_Improvedtreatmentdark, Bell:2023ysh_dec_ThermalizationAnnihilationDark}. These papers can be thought of as a series of works, with the original motivation credited to Nicole Bell, Giorgio Busoni and Sandra Robles. Given that they build off each other, there is significant overlap between each paper so that they are each a self-contained piece of work.

The chapters of this thesis are written in a way that combines the ideas presented in these papers in a logical manner, with much of the overlapping content removed. 
The chapters of this thesis and my contributions to these works are as follows:
\begin{itemize}
    \item Chapter~\ref{chapter:introduction} forms an original introduction and review of the literature.
    \item Chapter~\ref{chapter:compactobjects} is an original introduction to compact objects. The discussion of the white dwarf equation of state and observational status is in large part taken from section 2 of Ref.~\cite{Bell:2021fye_oct_Improvedtreatmentdark} that was written by me.
    \item Chapter~\ref{chapter:capture_intro} begins with an original review of Gould's formalism for dark matter capture in the Sun. The remainder of the chapter combines results from Refs.~\cite{Bell:2020jou_sep_ImprovedTreatmentDark, Bell:2020lmm_mar_ImprovedTreatmentDark} to give a complete overview of the capture formalism developed in these papers. For Ref.~\cite{Bell:2020jou_sep_ImprovedTreatmentDark}, I made contributions to developing the code used in the numerical calculations, with version 1 of the code written by Giorgio Busoni, and to the analytic calculations for the interaction rates. I contributed to writing the sections discussing details of the formalism and the numerical results. In Ref.~\cite{Bell:2020lmm_mar_ImprovedTreatmentDark}, I made similar contributions to the writing and numerical calculations as the previous paper.
    \item The first half of Chapter~\ref{chapter:capture_leptons} discusses the main results for the capture of dark matter due to scattering on the leptonic species in neutron stars from Ref.~\cite{Bell:2020lmm_mar_ImprovedTreatmentDark}. The second half looks at dark matter interaction with electrons in white dwarfs in Ref.~\cite{Bell:2021fye_oct_Improvedtreatmentdark}. In this paper, the calculations and discussions involving dark matter-ion interactions are the work of Maura Ramirez-Quezada and are not included in this thesis. The calculations and text discussing the dark matter-electron interactions are my own work. In addition, I wrote the discussion and performed the numerical calculations of the white dwarf structure and equation of state.
    \item Chapter~\ref{chapter:capture_baryons} is based on the results of Ref~\cite{Anzuini:2021lnv_nov_Improvedtreatmentdark}. I contributed to the numerical calculations of $m^*$ (relevant for multiple scattering) and deep inelastic scattering. Much of the numerical code was reused from the previous papers. I also contributed to writing the sections discussing the interaction rates and the results of the capture rates and threshold cross-sections. 
    \item Chapter~\ref{chapter:thermalisation} is based on Ref.~\cite{Bell:2023ysh_dec_ThermalizationAnnihilationDark}. I performed all numerical calculations and wrote the first draft of the paper, save section 2, which was written by Sandra Robles. I contributed to the analytical approximations together with Giorgio Busoni.
    \item Chapter~\ref{chapter:conclusion} presents my concluding remarks on this thesis as a whole and the future directions of this work.
\end{itemize}
The plots presented in the published versions of the papers were all made by Sandra Robles. Only when necessary were these figures remade before being included in this thesis.
Any part of these papers I did not originally write was rewritten before being included in this thesis. 

The results of Ref.~\cite{Bell:2020obw_sep_NucleonStructureStrong} were not used in this thesis as I did not make significant contributions to the text.

Ref.~\cite{Bell:2024qmj_apr_HeavyDarkMatter} was written in large part by Sandra Robles and Giorgio Busoni and, as such, is not included in this thesis either. However, in this work, I performed the numerical calculations for the dark matter thermalisation times and wrote the first draft of that discussion.
\end{preface}
